Synthesis of Intrusion Detection Rules \\ for Increased Network Security

### 2019-06-27 paper/paper.tex
quick pass on text

Network Intrusion Detection Systems (\nids{}) are an important
instrument that system administrators use to secure networks. These
systems monitor the network traffic for malicious behavior and act
accordingly by blocking messages or alerting humans about suspicious
events. \nids{} are typically placed behind a firewall, vetting the
traffic that the firewall did not block. These systems are very
popular today, with various public (\eg{}, Snort~\Fix{cite} and
Suricata~\Fix{cite}) and private implementations (\eg{}, \Fix{...} and
\Fix{...}). \Fix{...elaborate...}

\sloppy \nids{} can be categorized in two groups: 1)~rule-based and
2)~anomaly-based. A rule-based \nids{} checks if the network traffic
matches a given set of rules. A rule describes a potentially
mallicious message pattern. Figure~\Fix{small example} shows an
example rule of Suricata~\cite{suricata}, a popular open-source
\nids{}, owned and supported by the Open Information Security
Foundation (OISF)~\cite{oisf}. \Fix{quickly explain example}
Rule-based \nids{} focuses on known attacks whereas anomaly-based
\nids{} focuses on unanticipated issues. The hypothesis is that attack
manifestations produces traffic that deviates from regular benign
traffic that has been observed by the network. Anomaly-based \nids{}
have shown useful specially for detecting denial of service attacks
when determining the right threshold for the number of accesses that
triggers an alarm is challenging. Conceptually, these two approaches
are complementary. This paper focuses on rule-based \nids{}, which are
very popular both in academia and industry.

@@

Mantaining \nids{} rules can be a daunting activity. Although a set of
rules are typically provided by the \nids{} (\eg{}, \suri{} provides
\numrulessuri{} rules), 1) enabling all possible rules is not a
practical option and 2) rules not in the original rule set may need to
be added. The first statement is consequential of the inherent
physical limitation of rule-based \nids{}. The cost of checking rules
is proportional to the traffic load and the number of enabled
rules. If system capacity is insufficient to support the load, the
system can potentially miss packets and attacks\Fix{I guess we have
  some reference for this.}. To cope with that issue, system
administrators restrict the attack surface before applying
\nids{}---they block certain protocols on the firewall and configure
the \nids{} to use a selection of rules that apply to the authorized
traffic. The second problem is associated with the observation that
\numundetected{} of the attacks produced by
Metasploit~\cite{metasploit}, a popular open-source penetration
testing tool, went undetected by the \suri{} \nids{}. For example,
\Fix{...}. (Section~\ref{sec:suri-metas-coverage} details the coverage
of \suri{} against the exploits produced by \metas.) Unfortunately,
manually creating those rules can be time-consuming and
error-prone. Figure~\Fix{...} shows an example rule illustrating the
complexity of the Suricata rule language\footnote{Suricata extends the
  language used by Snort~\cite{snort}.}. Collectively, these issues
indicate that synthesizing \nids{} rules is an important aid for
system administrators.

>>>>

Network Intrusion Detection Systems (\nids{}) are an important
instrument that system administrators have to secure networks. These
software systems monitor the network traffic for malicious behavior
and act accordingly by blocking messages or alerting humans about
suspicious events. \nids{} are typically placed behind a firewall,
vetting the traffic that the firewall did not block. \nids{} are
frequently adopted today. Various open-source (\eg{}, Snort~\Fix{cite}
and Suricata~\Fix{cite}) and commercial implementations (\eg{},
\Fix{...}  and \Fix{...}) exist. \Fix{...elaborate...}

\sloppy \nids{} can be categorized in two groups: 1)~rule-based and
2)~anomaly-based. A rule-based \nids{} checks if the network traffic
matches a given set of rules. A rule describes a potentially
mallicious message pattern. Figure~\Fix{small example} shows an
example rule of Suricata~\cite{suricata}, a popular open-source
\nids{}, maintained by the Open Information Security Foundation
(OISF)~\cite{oisf}. \Fix{quickly explain example} Rule-based \nids{}
focuses on known attacks whereas anomaly-based \nids{} focuses on
unanticipated issues. The hypothesis of anomaly-based \nids\ is that
attack manifestations produces traffic that deviates from regular
benign traffic that has been already observed by the network and that
deviation is noticeable. Anomaly-based \nids{}, specially those that
use machine learning, have shown useful specially for detecting denial
of service attacks~\Fix{cite cite}. In that scenario, determining the
right threshold on the rate/volume of network accesses that should
trigger an alarm is challenging. Conceptually, these two approaches
are complementary. This paper focuses on rule-based \nids{}, which are
already very popular both in industry and academia. Our analysis
focuses on Suricata, but it applies to any other rule-based \nids{}.

@@

Mantaining \nids{} rules can be a daunting activity. Although the
\nids{} typically provides a set of pre-defined rules, such ruleset is
neither necessary nor sufficient. More precisely:

\begin{enumerate}
\item enabling all possible rules is not practical;
\item rules not in the original rule set may need to be added in a
  given setup.
\end{enumerate}

The first statement is consequential of the inherent physical
limitation of rule-based \nids{}. The cost of checking rules is
proportional to the traffic load and the number of enabled rules. If
system capacity is insufficient to support the load, the system can
potentially miss packets and attacks\Fix{I guess we have some
  reference for this.}. To cope with that issue, system administrators
restrict the attack surface before applying \nids{}---they block
certain protocols on the firewall and configure the \nids{} to use a
selection of rules that apply to the authorized traffic. The second
statement is supported by the observation that \numundetected{} of the
attacks produced by Metasploit~\cite{metasploit}, a popular
open-source penetration testing tool, went undetected by \suri{}. For
example, \Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides
more details.) Unfortunately, manually creating those rules from
scratch can be time-consuming and error-prone. Figure~\Fix{...}  shows
an example rule illustrating the complexity of the Suricata rule
language\footnote{Suricata extends the language used by
  Snort~\cite{snort}.}.  In addition, looking for rules from untrusted
sources (say, the internet) is unreliable. Collectively, these issues
suggest that synthesizing \nids{} rules is an important aid for system
administrators.

\Fix{why is that challenging (worth of research)?}
